---
title: "How does the daily occupancy rate of ON shelters change under the strike of COVID-19?"
author: "JIN QIU"
date: 2020-09-23T21:13:14-05:00
categories: ["R"]
tags: ["R Markdown", "plot", "regression"]
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
install.packages("opendatatoronto", repos="http://cran.us.r-project.org" )
install.packages("ggplot2", repos="http://cran.us.r-project.org" )
install.packages("tidyverse", repos="http://cran.us.r-project.org" )
install.packages("ggpubr", repos="http://cran.us.r-project.org" )
installed.packages("gridExtra", repos="http://cran.us.r-project.org" )
```
    
    
```{r, include=FALSE}  
install.packages("ggpubr")
```
# Abstract

This research aims at discovering the change of the situation of daily
shelters in ON due to the strike of COVID-19. The analysis is based on
the comparison between the data from year 2020 and year 2019, and will
especially focus on the changes in occupancy rate. The most important
finding is that the average occupancy rate has decreased, while the
situation of overwhelmed shelter is even more severe.

# Introduction

In year 2020, due to the strike of COVID-19, the Ontario government has
required everyone to keep the social distance and reduce the frequency
of going out to stop the spread of COVID-19. That means, all the
homeless people on the street has to go to a shelter. That makes me get
curious: will there be more shelters in Ontario to host all the homeless
people compared to last year? Are there enough spots in those shelters?
    To answer the questions, I did a statistical research and find out that
there might not be enough shelters in Ontario.

From the analysis, I find out that the number of shelters that are under
performance this year is even less than the number of last year(2019),
along with a decrease in the occupancy rate. The average occupancy rate
in 2019 is 0.9559, while that in 2020 is 0.7981. ALthough the situation
of a overwhelmed shelter has existed since last year, the highest
occupancy rate in ON this year for a single shelter has arrived 4.05,
which is almost 3 times of that in last year. This means, the use ratio
of the shelters is below the average during the period of COVID-19, and
there are not enough available spots provided in the shelter.

During the process of the analyzing the occupancy rate, we are using the
data from OPENDATA Toronto. And they collect the information of all the
shelters in Ontario everyday at 4 a.m. However, since I am doing the
research in the middle of 2020, the data I am using only covered that
past 9 months, that may cause some mistake when I use the data to do the
comparison and analysis. For example, there might be more homeless
people in the shelter in winter due to the weather, and the average
ratio might be underestimated with the use of current data. Another
possible error is, when I am doing the research, neither the data from
2020 or the data from 2019 is selected randomly, I literally used all
the data points that I am provided with. The conclusion that I can draw
from the current data is there is not enough spots in the shelters of ON
for the past 9 months, and the government should invest more to make
sure that all the homeless can stay safe in a shelter during this
special period . I will continue to do the research and decide how many
more shelters and spots should be provided in the future.

```{r,include=FALSE,echo=FALSE}
library(opendatatoronto)
library(dplyr)


package <- show_package("8a6eceb2-821b-4961-a29d-758f3087732d")
package


resources <- list_package_resources("8a6eceb2-821b-4961-a29d-758f3087732d")

datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))

data20 <- filter(datastore_resources, row_number()==1) %>% get_resource()
data20


data19 <- filter(datastore_resources, row_number()==2) %>% get_resource()
data19
```




# How does the sector that the shelter hosts change due to the impact of COVID-19 for year 2020 compared to year 2019?

```{r sector bar,echo=FALSE, message=FALSE}
library(ggplot2)
attach(data19)
hist19=ggplot(data19,aes(x= SECTOR))+
    geom_bar()+
    ylim(0,12500)+
    ggtitle("histogram for 2019")
detach(data19)
attach(data20)
hist20=ggplot(data20,aes(x= SECTOR))+
    geom_bar()+
    ylim(0,12500)+
    ggtitle("histogram for 2020")
library(ggpubr)
ggarrange(hist19,hist20,ncol =2,nrow=1)



```

From the plot above, we can conclude that the shelter provided for all
sectors has decreased this year more or less. There might be various
reasons for this situation, one of the most possible reason is that
COVID-19 has made it difficult for the government to hold all the
shelters on running due to the shortage of funding or workforce.


# Futhermore, how does the occupancy rate change due to the COVID-19 in year 2020 compared to year 2019?

```{r plots, echo=FALSE}
ratio19= data19$OCCUPANCY/data19$CAPACITY
ratio20= data20$OCCUPANCY/data20$CAPACITY
par(mfrow=c(1,2))
plot(data19$CAPACITY,data19$OCCUPANCY)
lmod19=lm(data19$OCCUPANCY~data19$CAPACITY)
abline(lmod19)
plot(data20$CAPACITY,data20$OCCUPANCY)
lmod20=lm(data20$OCCUPANCY~data20$CAPACITY)
abline(lmod20)

```

This two plots show that there might be some postive linear relationship
between capacity and occupancy. We can conclude that the variance of the
data in 2020 is larger than that of 2019. This means, the situations may
vary from diffrent shelters in 2020.

```{r}
summary(ratio19)
summary(ratio20)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
hist(ratio19)
hist(ratio20)
```

From the chart and the histograms above, it is not hard to conclude that
the average occupancy rate in year 2020 is about 20% lower than that in
2020. In specific, the histogram for year 2020 is hardly right-skewed,
while we could say the histogram for year 2019 is slightly right-skewed.
Although they both peak at a occupancy ratio=1, there are more extreme
data points in year 2020 compared to year 2019. This could lead to a
conclusion that although the shelters almost reach their maximum
capacity in both year and get overwhelmed to some extent, the situation
in 2020 may be more extreme, since there are more outliers.

In general, the situation we are facing right now is some shelters are
not fully used, while some others are extremely overloaded. The possible
reason for this might be that people can not commute conveniently due to
the COVID-19, so they are forced to stay in the certain area. However,
even if we are only looking at the data in year 2019, we cans say that
there is no enough shelter spots in ON, to fulfill the society’s needs
for shelter spots, the government needs to make more investment.










# References

1.  JJ Allaire and Yihui Xie and Jonathan McPherson and Javier Luraschi
and Kevin Ushey and Aron Atkins and Hadley Wickham and Joe Cheng and
Winston Chang and Richard Iannone (2020). rmarkdown: Dynamic
Documents for R. R package version 2.1. URL
<https://rmarkdown.rstudio.com>.

Yihui Xie and J.J. Allaire and Garrett Grolemund (2018). R Markdown: The
Definitive Guide. Chapman and Hall/CRC. ISBN 9781138359338. URL
<https://bookdown.org/yihui/rmarkdown>.

2.  Wickham et al., (2019). Welcome to the tidyverse. Journal of Open
Source Software, 4(43), 1686, <https://doi.org/10.21105/joss.01686\>
    
    3.  H. Wickham. ggplot2: Elegant Graphics for Data Analysis.
Springer-Verlag New York, 2016.

4. Baptiste Auguie (2017). gridExtra: Miscellaneous Functions for “Grid”
Graphics. R package version 2.3.
<https://CRAN.R-project.org/package=gridExtra>
    
    5.  Alboukadel Kassambara (2020). ggpubr: ‘ggplot2’ Based Publication
Ready Plots. R package version 0.4.0.
<https://CRAN.R-project.org/package=ggpubr>
    
    6.  Open Data Dataset. (2020). Retrieved September 23, 2020, from
<https://open.toronto.ca/dataset/daily-shelter-occupancy/>
